---
title: "BA64060_Assignment4"
author: "Durga Prasad Gandi"
date: "2023-11-10"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load the required packages
```{r}
library(flexclust)
```
```{r}
library(cluster)
library(tidyverse) 

```
```{r}

library(factoextra)
```
**This library is used to extract and visualize information from the results of multivariate data analyses**

```{r}
library(FactoMineR)

library(ggcorrplot)
```
## Load the dataset

```{r}
pharmaceutical = read.csv("C:/Users/gdurg/Downloads/Pharmaceuticals.csv")
pharmaceutical1 = pharmaceutical[3:11]
head(pharmaceutical1)
```

First entity has the highest market cap at 68.44, while 3rd has the lowest at 6.30.Market cap is a measure of a company's size, calculated by doing product of stock price by the number of outstanding shares

With a beta of 1.11, entity 6 has the highest volatility when compared to the market.**

With a PE ratio of 82.5, Entity 2 has the highest, indicating a rather expensive valuation.
The PE ratio, which shows how much investors are ready to pay for each dollar of earnings, is calculated by dividing the company's stock price by its earnings per share.

With the lowest asset turnover (0.6), Entity 5 may be using its assets less effectively to drive sales.
The effectiveness with which a business generates revenue from its assets is gauged by asset turnover.

With a leverage ratio of 0.60, Entity 2 has the highest level of debt in its capital structure.
Leverage quantifies how much debt a business uses to fund its operations.

5th Entity has experienced a significant rise in sales, as evidenced by its highest revenue growth of 26.81%.
The percentage rise in revenues over a given period for a corporation is known as revenue growth.

With a net profit margin of 16.1%, Entity 1 has the largest, suggesting a comparatively greater percentage of revenue kept as profit.

**These measures provide information on a number of financial performance factors for each organization, such as leverage, profitability, valuation, and effective use of assets. Additionally, the industry and unique circumstances of each business may influence interpretation.**

```{r}
summary(pharmaceutical1)
```

These organizations have market capitalizations that range widely, from 0.41 to 199.47. With a median market cap of 48.19 and an average of 57.65, the distribution appears to be slightly positively biased. With a market capitalization of 68.44, Entity 1 has the largest, and Entity 6 has the lowest, at 16.90.

A stock's beta, or volatility relative to the market, can be anywhere between 0.18 and 1.11. With Entity 6 having the highest volatility at 1.11, the average beta is 0.5257. With a median beta of 0.46, most entities have a central tendency around this value.

The PE ratios show that there is a wide range between 3.60 and 82.50. With a median of 21.50, the mean PE ratio is 25.46. The significant discrepancy between the mean and median points to the possibility of higher-order outliers affecting the mean.

A company's profitability as a percentage of shareholder equity, or ROE, can range from 3.9% to 62.9%. With a mean ROE of 25.8%, the average return is large. The distribution looks dispersed, though, with Entity 4 having the highest ROE (27.3).

Growth rates range from -3.17% to 34.21%, indicating varied revenue trajectories.
Entity 5 experiences the highest growth at 26.81%.

ROE varies from 3.9% to 62.9%, indicating diverse profitability.
Entity 4 stands out with the highest ROE at 27.4, showcasing effective use of equity.

Net profit margin ranges from 2.6% to 25.5%, representing profitability.
Entity 1 leads with the highest margin at 16.1%.

**The information provides a thorough understanding of the financial dynamics of these companies, including market value, risk exposure, valuation, profitability, effective use of assets, and potential for growth. In order to understand the complex financial profiles of each organization, interpretation should take into consideration both central tendencies and the distribution of values.**




```{r}
pharmaceutical2 <- scale(pharmaceutical1)
row.names(pharmaceutical2) <- pharmaceutical[,1]
distance <- get_dist(pharmaceutical2)
corr <- cor(pharmaceutical2)
fviz_nbclust(pharmaceutical2, kmeans, method = "silhouette")
```

# The cluster k=5 is the highest cluster with having average of silhouette width more than 0.2 


```{r}
set.seed(1)
k5 <- kmeans(pharmaceutical2, centers = 5, nstart = 25) # k = 5 i.e number of restarts =25
k5$centers 

```
```{r}
k5$size
```

```{r}
fviz_cluster(k5,data = pharmaceutical2)
```
**There are five different clusters where cluster 5 is below the average for both dim1 and dim2 .Cluster 3,4 is almost same as cluster5 but former(3,4) has more in dim2 than cluster5.Cluster 2 is more than average for dim1 but less than average for dim2,Interestingly Cluster 1 has both dim1 and dim2 values higher than average.**



```{r}
fviz_nbclust(pharmaceutical2, kmeans, method = "wss")
```
**Observing the chart, we can say that the elbow point 4 offers the optimal value for k. WSS will decrease with increasing k values, but we must weigh the trade-off between over fitting—a model that fits both noise and signal—and bias in the model. In this particular instance,The elbow point offers that middle ground where WSS can be reduced without going below reduces significantly more slowly when k = 4. Stated otherwise, k=4 yields the optimal value for bias and over fitting.**



```{r}
set.seed(35)
k51 = kcca(pharmaceutical2, k=5, kccaFamily("kmedians"))
k51
```

**set.seed is used to get the same cluster for same code if we run and k medians is used to determine median rather than means ,here Cluster 4 is the largest, with 7 observations, followed by Cluster 5 with 6observations.Cluster 2 is the smallest, with only 1 observation.**

```{r}
clusters_index <- predict(k51)
dist(k51@centers)
```
Distance function is representing the pairwise distances between cluster centers.if  the distance is smaller for cluster then it indicate they are closer,Here distance between cluster 2 and cluster4 is approx 3.12


```{r}
image(k51)
points(pharmaceutical2, col=clusters_index, pch=19, cex=0.3)
```

```{r}
pharmaceutical1 %>% mutate(Cluster = k5$cluster) %>% group_by(Cluster) %>%
summarise_all("mean")
```
The above results useful for Investors or analysts understanding the average financial characteristics of companies within each cluster.
For example, Cluster 5 has the highest average Market Cap, ROE, ROA, and Asset Turnover, suggesting that companies in this cluster may be considered high-performing and have a strong market position.
On the other hand, Cluster 4 has a relatively low Market Cap and high Leverage, indicating potential financial risk.



```{r}
clusplot(pharmaceutical2,k5$cluster, main="Clusters",color = TRUE, labels = 3,lines =0)
```
The plot describes the data points in a two-dimensional, with each point representing an observation  and its position are determined by the variables in your data set .Points belonging to the same cluster are likely to have the same color, making it easy to visually identify clusters.

### Question3

```{r}
pharmaceutical3 <- pharmaceutical[12:14] %>% mutate(Clusters=k5$cluster)
ggplot(pharmaceutical3, mapping = aes(factor(Clusters), fill
=Median_Recommendation))+geom_bar(position='dodge')+labs(x ='Clusters')
ggplot(pharmaceutical3, mapping = aes(factor(Clusters),fill =
Location))+geom_bar(position = 'dodge')+labs(x ='Clusters')
```

Here the Cluster3 which has median (hold) of greater values compared to all clusters
Cluster 1,2 both have moderate buy count,and for cluster1  moderate buy=moderate sell.
For 5th cluster and cluster2  moderate buy=hold but cluster5 has more count than 5th cluster.
We can observe that only cluster3 has most of the recommendations,Strong buy count whereas other cluster does not have,also same count as cluster2



```{r}
ggplot(pharmaceutical3, mapping = aes(factor(Clusters),fill =
Exchange))+geom_bar(position = 'dodge')+labs(x ='Clusters')
```




The resulting plot is a grouped bar plot where each cluster (Clusters) is represented on the x-axis as a separate group, and the bars within each group are colored based on the Exchange variable.
Overall comparing from chart Cluster 3 has highest exchange count,where as 1 and 5 has same cluster.
Cluster4 has all types of exchange rates but of smaller count
Cluster2 has more count than cluster4 but less than the remaining exchanges types.
Each bar graph represents the frequency or count of observations comes under into a specific combination of cluster and exchange.It helps in visualizing the distribution of observations across different clusters and exchanges. Plot above we can say that they are  useful when you want to compare the distribution of a categorical variable (Exchange) across different levels of another categorical variable (Clusters).
plot gives us insights into how the distribution of exchanges changes with in clusters.
