---
title: "Qmm_Assignment5"
author: "Durga Prasad Gandi"
date: "2023-11-04"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



#### The Research and Development Division of the Emax Corporation has developed three new products. A decision now needs to be made on which mix of these products should be produced. Management wants primary consideration given to three factors: total profit, stability in the workforce, and achieving an increase in the company’s earnings next year from the $60 million achieved this year. In particular, using the units given in the following table, they want to Maximize Z = P - 5C - 2D, where P = total (discounted) profit over the life of the new products, C = change (in either direction) in the current level of employment, D = decrease (if any) in next year’s earnings from the current year’s level. The amount of any increase in earnings does not enter into Z, because management is concerned primarily with just achieving some increase to keep the stockholders happy. (It has mixed feelings about a large increase that then would be difficult to surpass in subsequent years.)


```{r}
library(kableExtra)
```


```{r}
data_res <- data.frame(
Factor= c("Total profit", "Employment level", "Earnings next year"),
                   "1"= c(15,8,6),
                   "2"=c(12,6,5),
                   "3"=c(20,5,4),
                   Goal=c("Maximize","=70", ">=60"),
                   Units=c("Millions of dollars", "Hundreds of dollars", "Millions of dollars"))
data_res
```

```{r}
data_res %>%
 kable(align = "c") %>%
 kable_classic() %>%
 add_header_above(header = c(" "=1,"Product"=3," "=2)) %>%
 add_header_above(header = c(" "=1,"Unit contribution"=3," "=2)) %>%
 column_spec(1,border_right = TRUE) %>%
 column_spec(4,border_right = TRUE) %>%
 column_spec(5,border_right = TRUE)
```

#### Define y1+ and y1-, respectively, as the amount over (if any) and the amount under (if any) the employment level goal. Define y2+ and y2- in the same way for the goal regarding earnings next year. Define x1, x2, and x3 as the production rates of Products 1, 2, and 3, respectively. With these definitions, use the goal programming technique to express y1+, y1- , y2+ and y2- algebraically in terms of x1, x2, and x3. Also, express P in terms of x1, x2, and x3.

Let us assume $x_1$= Product 1 production rate

   $x_2$=  Product 2 production rate  
     
  $x_3$= Product 3 production rate

*Production rates of the Products are x1,x2 and x3 and the Employment level goal,goal for Earnings next year are considered as Constraints, so there is no possibility of conversion of constraints to Products x1,x2 and x3.*

The constraints are as follows:

constraint for Employment level:
$$8x_1+6x_2+5x_3=70$$

Constraint for Earnings for Next Year:
$$6x_1+5x_2+4x_3≥60$$

Define for suppose 
$$y_i=y_i^{−}− y_i^{+}$$

$y_1$= Represents Change (in either direction) in the current level of employment (C) $$y_1=y_1^{−}− y_1^{+}$$

$y_2$= Decrease (if any) in next year’s earnings from the current year’s level(D) 
 $$y_2=y_2^{−}− y_2^{+}$$

Where, 
$y_1^{+}$ is positive deviation or over achievement of goal for employment level. 
$y_1^{-}$ = negative deviation or under achievement of goal for employmentlevel. 
$y_2^{+}$ =positive deviation or over achievement of goal for Earnings next year.
$y_2^{-}$ = negative deviation or under achievement of goal for Earnings next year.

Constraints in terms of deviation form can be represented as $y_{1}, y_{2}$:

$$y_{1}=8x_1+6x_2+5x_3−70$$
$$y_1^{+}−y_1^{-}=8x_1+6x_2+5x_3−70$$

$$y_1^{+}=8x_1+6x_2+5x_3−70+y_1^{-}$$

$$y_1^{-}=y_1^{+}−(8x_1+6x_2+5x_3−70)$$

$$y_2=6x_1+5x_2+4x_3−60$$

$$y_2^{+}−y_2^{-}=6x_1+5x_2+4x_3−60$$


$$y_2^{+}=6x_1+5x_2+4x_3−60+y_2^{-}$$


$$y_2^{-}=y_2^{+}−(6x_1+5x_2+4x_3−60)$$

Express P in terms of x1, x2, and x3

$$P=15x_1+12x_2+20x_3$$

Express management’s objective function in terms of $x_1, x_2, x_3, y_1^+, y_1^- , y_2^+ ,y_2^-$

OBJECTIVE FUNCTION:
$$MaximizeZ=P−5C−2D$$

P = total (discounted) profit over the life of the new products $$P=15x_1+12x_2+20x_3$$

C = change (in either direction) in the current level of employment $$y_1=y_1^{+}−y_1^{-}$$

D = decrease (if any) in next year’s earnings from the current year’s level $$y_2=y_2^{+}−y_2^{-}$$


Maximize Z=P−5C−2D

$$MaxZ=P−5y_1−2y_2$$

*Expressing management’s objective function in terms of x1, x2, x3, y1+, y1- , y2+ and y2-*

$$MaxZ = 15x_1+12x_2+20x_3−5(y_1^{-}+y_1^{+})-2(y_2^{-})-0(y_2^{+})$$

#### 3.Formulate and solve the linear programming model. What are your findings?


```{r}
library(lpSolve)
library(lpSolveAPI)
```


```{r}
solveob <- make.lp(2, 7)
```
*Here Two constraints and Seven decision variables of linear programming will be created.*

```{r}
set.objfn(solveob, c(15,12,20,-5,-5,0,-2))
```

*solveob is an object that identifies the LP problem it will then create the objective function coefficients for the Seven decision variables.The coefficients are provided as a vector in this scenario. It is important to keep the variables in their correct arrangement. Constraints and the objective function should follow a consistent order.*


```{r}
lp.control(solveob, sense = 'max')
```
*The objective sense will subject  to ‘max’ suggests that objective function is to be  maximized .*

```{r}
set.row(solveob,1,c(8,6,5,-1,1,0,0), indices = c(1,2,3,4,5,6,7))
```
*First Constraint for LP problem* 

```{r}
set.row(solveob,2,c(6,5,4,0,0,-1,1), indices = c(1,2,3,4,5,6,7))
```
*This will add the second constraint to the LP problem*


```{r}
rhs<-c(70, 60)
set.rhs(solveob,rhs)
```
*rhs is represented as right-hand side values for the constraints in the LP problem.*

```{r}
set.constr.type(solveob,c("=","="))
set.bounds(solveob,lower = rep(0,7))
```
*It will then subject the constraint types for the three constraints.Here, they are all set to equality (‘=’).*

```{r}
lp.rownames<-c("employment","investment")
```
*This assigns names to the two constraintsi.e employment and investment.*


```{r}
lp.colnames<-c("x1","x2","x3","y1p","y1m","y2p","y2m","y3p","y3m")
```
*It assigns names to the nine decision variables. p stands for positive and m for negative.*


```{r}
solve(solveob)
```
*This line solves the LP problem using the specified constraints and objective function.*

```{r}
get.objective(solveob)
```

```{r}
get.variables(solveob)
```

#### Findings:

**These findings demonstrate the goal programming problem's ideal solution. Considering the specified limitations and penalties, the values of the decision variables give information about the suggested production levels and objective deviations for each factor. This tactic aims to optimize profit by deducting the target deviations and accounting for the associated penalties.The optimal solution to the LP problem is 275. this shows that we have achieved the best outcome possible given the constraints, with an objective function value of 275.**
**The choice variables define the best option, but the slack variables demonstrate that the restrictions are either fully satisfied or have a surplus.**

Decision variable values are:
$x_1=0$
$x_2=0$
$x_3=15$
$y_1^+=5$
$y_1^-=0$
$y_2^+=0$
$y_2^-=0$

***Given that $y_1^+$ only seems to have a non-zero value, the employment goal appears to have a positive variance. To calculate the penalty associated with, you need to know the related penalty coefficient, which according to the code is 5. As a result, the penalty for the positive employment deviation is as follows: 5*5= 25 There is no penalty for the other factors() related variables because their values are zero in the optimal solution***

